---
title: "Coding Practice 7"
author: "Mary Fantasia"
date: "2026-03-31"
output:
  html_document:
    toc: true
    toc_float: true
  md_document:
    variant: gfm
  pdf_document:
    toc: true
---

# load the packages
```{r load packages}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
```

# Load data
```{r load data}
data("mtcars")
```

## Plot data
```{r plot data}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_smooth(method = lm, se = FALSE, color = "grey") +
  geom_point(aes(color = wt)) +
  xlab("Weight") + 
  ylab("Miles per gallon") +
  scale_colour_gradient(low = "forestgreen", high = "black") +
  theme_classic()
```

### Run a linear model
```{r linear model}
lm(mpg~wt, data = mtcars)
```
### summary of linear model
```{r summary}
summary(lm(mpg~wt, data = mtcars))
```

### ANOVA
```{r anova}
anova(lm(mpg~wt, data = mtcars))
```

### test for correlation
```{r}
cor.test(mtcars$wt, mtcars$mpg)
```
## Plot residuals from linear model
```{r}
model <- lm(mpg~wt, data = mtcars)

ggplot(model, aes(y = .resid, x = .fitted)) +
  geom_point() +
  geom_hline(yintercept = 0)
```

# New data
```{r}
bull.rich <- read.csv("Bull_richness.csv")
```

### plot data
```{r}
bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.") %>%
  ggplot(aes(x = Fungicide, y = richness)) + 
  geom_boxplot()
```

### run a t-test
```{r}
bull.rich.sub <- bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.")

t.test(richness~Fungicide, data = bull.rich.sub)
```
### plug into linear model to get the same results
```{r}
t.test(richness~Fungicide, data = bull.rich.sub, var.equal = TRUE)
```
```{r}
summary(lm(richness~Fungicide, data = bull.rich.sub))
```

```{r}
anova(lm(richness~Fungicide, data = bull.rich.sub))
```

### filter our data by growth stage
```{r}
bull.rich.sub2 <- bull.rich %>%
  filter(Fungicide == "C" & Treatment == "Conv." & Crop == "Corn")
```

### and visualize
```{r}
bull.rich.sub2$GrowthStage <- factor(bull.rich.sub2$GrowthStage, levels = c("V6", "V8", "V15"))

ggplot(bull.rich.sub2, aes(x = GrowthStage, y = richness)) +
  geom_boxplot()
```

### run different models to get the same result
```{r}
lm.growth <- lm(richness ~ GrowthStage, data = bull.rich.sub2)
summary(lm.growth)
```
```{r}
anova(lm.growth)
```
```{r}
summary(aov(richness ~ GrowthStage, data = bull.rich.sub2))
```

### post-hoc tests to find differences
```{r}
lsmeans <- emmeans(lm.growth, ~GrowthStage) # estimate lsmeans of variety within siteXyear
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey ajustment by default. 
Results_lsmeans
```
```{r have to install this package before that can run}
library(multcompView)
```
### filter data to include term
```{r}
bull.rich.sub3 <- bull.rich %>%
  filter(Treatment == "Conv." & Crop == "Corn")

bull.rich.sub3$GrowthStage <- factor(bull.rich.sub3$GrowthStage, levels = c("V6", "V8", "V15"))
```

### set up linear models
```{r}
# write it like this
lm.inter <- lm(richness ~ GrowthStage + Fungicide + GrowthStage:Fungicide, data = bull.rich.sub3)

# or like this
lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3)
```

### look at overlap
```{r}
summary(lm.inter) # significant terms
```
```{r}
anova(lm.inter) # The interaction term is signifant. 
```
### plot the effect of fungicide over the levels of growthstage
```{r}
bull.rich.sub3 %>%
  ggplot(aes(x = GrowthStage, y = richness, fill = Fungicide)) +
  geom_boxplot()
```
### and post-hoc test
```{r}
lsmeans <- emmeans(lm.inter, ~Fungicide|GrowthStage) # estimate lsmeans of variety within siteXyear
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeans
```
# Mixed effects models
lets use the interaction model from before
```{r}
lme0 <- lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3)
```

```{r}
lme1 <- lmer(richness ~ GrowthStage*Fungicide + (1|Rep), data = bull.rich.sub3)
summary(lme1)
```
```{r}
summary(lme0)
```
```{r}
summary(lme1)
```
For some terms, it actually makes the standard error of our betas go down, which means we have better estimates of our betas, and we are more confident that we are estimating them correctly in our linear models, which means we have more power to detect differences.

