---
title: "Coding Practice 5"
author: "Mary Fantasia"
date: "2026-03-05"
output: 
  md_document:
    variant: gfm
  html_document: 
   toc: true  
   toc_float: true  
---

#Data Wrangling
##First step, install R package, "tidyverse" then load the package

```{r}
install.packages("tidyverse")

library(tidyverse)
```

Loading this package will also load all of the listed packages above, so you do not have to load them individually.

##Next, check out these functions:

```{r}
# Data wrangling & manipulation

mutate()
select()
filter()
the pipe %>%
summarise()
group_by()
joining
pivotting
Integration with plotting

```

##Tutorial

```{r}
# Load data
microbiome.fungi <- read.csv("Bull_richness.csv")
str(microbiome.fungi)
```
Here is our data above.

###select()
```{r}
# Selecting only certain parts of outr data
microbiome.fungi2 <- select(microbiome.fungi, SampleID, Crop, Compartment:Fungicide, richness)
```

###filter()
```{r}
# Subset data
head(filter(microbiome.fungi2, Treatment == "Conv."))
```
Above shows our subsetted data.

```{r}
# To be more complex using &
head(filter(microbiome.fungi2, Treatment == "Conv." & Fungicide == "C"))
```
Above shows our complex way or subsetting data

```{r}
# To be more cmplex using or, |
head(filter(microbiome.fungi2, Sample == "A" | Sample == "B")) # samples A or B
```
Above shows our more complex way of subsetting data.

###mutate()
```{r}
# Create a new column called logRich
head(mutate(microbiome.fungi2, logRich = log(richness)))
```
Here is our data with the new column added.

```{r}
# Creating a new column which combines Crop and Treatment 
head(mutate(microbiome.fungi2, Crop_Treatment = paste(Crop, Treatment)))
```
Here is our data with the new column that combines data

###%>%
```{r}
# The pipe
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  head() # displaying the first six rows
```
Here we do not have to specify functions, previous step to next step.

###summarise()
```{r}
# Find things like means, standard deviation, errors
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich)) # calculating overall mean log richness within the conventionally managed treatment
```
Here is the mean of this data.

```{r}
# Connect multiple summary stats
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
```
Here we found mean, standard deviation, and standard error.

###group_by()
```{r}
# Get summary statistics by group
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
```
Here we have statistics by group.

###Connecting to plotting
```{r}
# Group of functions work together nicely creating a nice plot
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  ggplot(aes(x = Fungicide, y = Mean.rich)) + # adding in a ggplot
  geom_bar(stat="identity") +
  geom_errorbar( aes(x=Fungicide, ymin=Mean.rich-std.err, ymax=Mean.rich+std.err), width=0.4) +
  theme_minimal() +
  xlab("") +
  ylab("Log Richness") +
  facet_wrap(~Treatment)
```
Here is our plot!

###Joining
```{r}
# selecting just the richness and sample ID
richness <- microbiome.fungi %>%
  select(SampleID, richness)

# selecting columns that don't include the richness
metadata <- microbiome.fungi %>% 
  select(SampleID, Fungicide, Crop, Compartment, GrowthStage, Treatment, Rep, Sample)

head(metadata)
```
Here are rows based on certain criteria, in this case, metadata.

```{r}
head(richness)
```
Here are rows based on richness.

```{r}
# adding the richness data to the metadata based on on the common column of sampleID
head(left_join(metadata, richness, by = "SampleID")) 
```
More joining based on certain criteria.

###pivot_wider()
```{r}
# Pivoting to convert from wide to long format and back again
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) # calculates the mean per Treatment and Fungicide 
```
Here we have our data ready to pivot.

```{r}
# Pivoting data
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) # pivot to wide format
```
Here is our pivoted data.

```{r}
#Taking the difference between the fungicide and control
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) # calculate the difference between the means. 
```
Here is our data ready to plot.

```{r}
# Plot difference
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) %>%  # calculate the difference between the means. 
  ggplot(aes(x = Treatment, y = diff.fungicide)) + # Plot it 
  geom_col() +
  theme_minimal() +
  xlab("") +
  ylab("Difference in average species richness")
```
Here is our new plot for our pivoted data.






