---
title: "Coding Challenge 5"
author: "Mary Fantasia"
date: "2026-03-12"
output: 
  html_document: 
   toc: true  
   toc_float: true  
  md_document:
    variant: gfm
---

# Question 1
Download two .csv files from Canvas called DiversityData.csv and Metadata.csv, and read
them into R using relative file paths.
```{r Read files using relative file paths}
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)

DiversityData <- read.csv("DiversityData.csv")
str(DiversityData)
Metadata <- read.csv("Metadata.csv")
str(Metadata)
```

# Question 2
Join the two dataframes together by the common column ‘Code’. Name the resulting dataframe
alpha.
```{r Join two dataframes}
alpha <- DiversityData %>%
  left_join(Metadata, by = "Code")
```

# Question 3
## Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by the Shannon diversity index (column Shannon) divided by the log of the richness column.
A. Using mutate, create a new column to calculate Pielou’s evenness index.
B. Name the resulting dataframe alpha_even.
```{r Creating a new column}
head(mutate(alpha, alpha_even = log(richness)))
```
# Question 4
## Using tidyverse language of functions and the pipe, use the summarise function and tell me the mean and standard error evenness grouped by crop over time.
A. Start with the alpha_even dataframe
B. Group the data: group the data by Crop and Time_Point.
C. Summarize the data: Calculate the mean, count, standard deviation, and standard error for the even variable within each group.
D. Name the resulting dataframe alpha_average
```{r Finding mean and standard error evenness}
alpha_average <- alpha %>%                              # THIS CODE WAS WRITTEN WITH HELP FROM COPILOT.
  select(Code, Crop, Time_Point, richness) %>%          # select just what you need
  mutate(alpha_even = log(richness)) %>%                # create the log-evenness column
  group_by(Crop, Time_Point) %>%                        # group by crop over time
  summarise(
    n         = sum(!is.na(alpha_even)),                # count of non-missing values
    mean_even = mean(alpha_even, na.rm = TRUE),         # mean
    sd_even   = sd(alpha_even, na.rm = TRUE),           # standard deviation
    se_even   = sd_even / sqrt(n),                      # standard error
    .groups   = "drop"
  )
alpha_average
```

# Question 5
## Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
A. Start with the alpha_average dataframe
B. Select relevant columns: select the columns Time_Point, Crop, and mean.even.
C. Reshape the data: Use the pivot_wider function to transform the data from long to wide
format, creating new columns for each Crop with values from mean.even.
D. Calculate differences: Create new columns named diff.cotton.even and diff.soybean.even
by calculating the difference between Soil and Cotton, and Soil and Soybean,
respectively.
E. Name the resulting dataframe alpha_average2
```{r Calculating differences}
alpha_average2 <- alpha_average %>% #THIS CODE WAS WRItTEN WITH HELP FROM COPILOT.
  # B) Select relevant columns (supports either mean_even or mean.even)
  select(Time_Point, Crop, any_of(c("mean_even", "mean.even"))) %>%
  # Standardize the mean column name to mean_even for the rest of the pipeline
  rename(mean_even = any_of(c("mean_even", "mean.even"))) %>%
  # C) Go from long (one row per Crop) to wide (one column per Crop)
  pivot_wider(
    names_from  = Crop,
    values_from = mean_even
  ) %>%
  # D) Compute differences (Soil − Cotton, Soil − Soybean)
  mutate(
    diff.cotton.even  = Soil - Cotton,
    diff.soybean.even = Soil - Soybean
  )
alpha_average2
```
# Question 6
## Connecting it to plots
A. Start with the alpha_average2 dataframe
B. Select relevant columns: select the columns Time_Point, diff.cotton.even, and
diff.soybean.even.
C. Reshape the data: Use the pivot_longer function to transform the data from wide to long
format, creating a new column named diff that contains the values from diff.cotton.even
and diff.soybean.even.
D. Create the plot: Use ggplot and geom_line() with ‘Time_Point’ on the x-axis, the column
‘values’ on the y-axis, and different colors for each ‘diff’ category. The column named
‘values’ come from the pivot_longer.
```{r Plotting the data}
# THIS CODE WAS WRITTEN WITH HELP FROM COPILOT
plot_data <- alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer(c(diff.cotton.even, diff.soybean.even),names_to = "diff", values_to ="values") %>%
  mutate(Time_Point = suppressWarnings(as.numeric(Time_Point)))
plot_data

ggplot(plot_data, aes(x = Time_Point, y = values, color = diff, group = diff)) +
  geom_line(linewidth = 1) +         # the requested line plot
  geom_point(size = 2) +             # optional: shows points at each time
  labs(
    x = "Time Point",
    y = "Difference in Mean Log Evenness",
    color = "Difference Type",
    title = "Differences in Evenness Over Time"
  )

```

# Github link
Here is a link to my Github where this is in the coding_challenge_5 directory

https://github.com/maryfantasia/Scientific_Reproducibility_Class


