---
title: "Coding Challenge 6"
author: "Mary Fantasia"
date: "2026-03-19"
output:
  html_document:
    toc: true
    toc_float: true
  md_document:
    variant: gfm
  pdf_document:
    toc: true
---
# Question 1
## Regarding reproducibility, what is the main point of writing your own functions and iterations? 
The main point of writing your own functions and iterations for reproducibility is to prevent copy/paste errors while writing many of the same lines of code.

# Question 2
## : In your own words, describe: 1) how to write a function in R, and 2) how to write a for loop in R.
1) To write a function in R, you insert a code into a variable.
2) To write a for loop in R, you give the code an "if, then" objective.

# Question 3
## Read the data
```{r read the data}
cities <- read.csv("Cities.csv")
```

# Question 4
## Write a function that calculates the distance (in kilometers) between two pairs of geographic coordinates using the Haversine formula. 
```{r write a function}
haversine_distance <- function(lat1, lon1, lat2, lon2) {
  # convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180

# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a)) 

# Earth's radius in kilometers
earth_radius <- 6378137

# Calculate the distance
distance_km <- (earth_radius * c)/1000
return(distance_km)
}
```

# Question 5
## . Using your function from question 4, calculate the distance between Auburn, AL and New York City.
```{r calculate distance}
### STEP 1 — Subset Auburn and NYC ###
auburn <- subset(cities, city == "Auburn")
nyc   <- subset(cities, city == "New York")

### STEP 2 — Extract the latitude/longitude values ###
lat1 <- auburn$lat
lon1 <- auburn$long
lat2 <- nyc$lat
lon2 <- nyc$long

### STEP 3 — Calculate the distance ###
distance_auburn_nyc <- haversine_distance(lat1, lon1, lat2, lon2)

### STEP 4 — Print the result ###
distance_auburn_nyc

```

# Question 6
## Use your distance function inside a for loop to calculate the distance between Auburn, AL and every other city in Cities.csv. Each iteration should calculate the distance for one city.
```{r auburn vs other cities}
city_all <- unique(cities$city)

null_df <- NULL
for (i in seq_along(city_all)){
  auburn_lat=auburn$lat
  auburn_long=auburn$long
  city_lat=cities$lat[cities$city == city_all[[i]]]
  city_long=cities$long[cities$city == city_all[[i]]]
  distance <- haversine_distance(auburn_lat, auburn_long, city_lat, city_long)
  
  distance_all <- data.frame(distance, city_all[[i]])
null_df <- rbind.data.frame(null_df, distance_all) 
  }
print(null_df)
```

# Question 7
## Modify your loop so that each iteration appends one new row to a dataframe with the following columns:

•	City1 (the other city)
•	City2 (Auburn)
•	Distance_km

```{r auburn vs all other cities}
city_all <- unique(cities$city)

# Empty dataframe with correct column names
distance_df <- data.frame(
  City1 = character(),
  City2 = character(),
  Distance_km = numeric(),
  stringsAsFactors = FALSE
)

for (i in seq_along(city_all)) {

  # Coordinates for Auburn (City2)
  auburn_lat  <- auburn$lat
  auburn_long <- auburn$long

  # Coordinates for the other city (City1)
  city_lat  <- cities$lat[cities$city == city_all[i]]
  city_long <- cities$long[cities$city == city_all[i]]

  # Calculate the distance
  distance_km <- haversine_distance(auburn_lat, auburn_long, city_lat, city_long)

  # Create one-row dataframe for this iteration
  new_row <- data.frame(
    City1 = city_all[i],
    City2 = "Auburn",
    Distance_km = distance_km,
    stringsAsFactors = FALSE
  )

  # Append
  distance_df <- rbind(distance_df, new_row)
}

print(distance_df)
```
*This code was written using help from copilot.
