---
title: "Coding challenge 7"
author: "Mary Fantasia"
date: "2026-04-01"
output:
  html_document:
    toc: true
    toc_float: true
  md_document:
    variant: gfm
  pdf_document:
    toc: true
---

# Coding Challenge 7

## Question 1
Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor

```{r Load libraries}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
library(ggplot2)
library(dplyr)
library(stringr)

```
```{r Read data}
plant.emergence <- read.csv("PlantEmergence.csv")
```

```{r Creating factors}
# Convert Treatment, DaysAfterPlanting, and Rep to factors
plant.emergence$Treatment <- as.factor(plant.emergence$Treatment)
plant.emergence$DaysAfterPlanting <- as.factor(plant.emergence$DaysAfterPlanting)
plant.emergence$Rep <- as.factor(plant.emergence$Rep)
```

## Question 2
Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 

```{r Fitting the linear model}
emergence_lm <- lm(Emergence ~ Treatment * DaysAfterPlanting, data = plant.emergence)
```

```{r View the model summary}
summary(emergence_lm)
```

```{r View ANOVA results}
anova(emergence_lm)
```

## Question 3
Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

### Interpretation of question 2
Based on the results of the linear mode in question 2, the interaction for Treatment and DaysAfterPlanting seperately are significant, but the Treatment:DaysAfterPlanting is not. Fitting a simplified model is justified.

```{r A simplified linear model}
# Fit a simplified linear model without the interaction term
emergence_lm_simple <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = plant.emergence)
```

```{r New model summary}
summary(emergence_lm_simple)
```

```{r View new ANOVA results}
anova(emergence_lm_simple)
```
### Interpretation of question 3
There is a major difference between Treatment 1 and Treatment 2, specifically 134.531 fewer emerging individuals in Treatment 2 than plants receiving Treatment 1, and this result is significant.

## Question 4
Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.

```{r Calculate least square means}
# Calculate least square means for Treatment
treat_means <- emmeans(emergence_lm_simple, ~ Treatment)
```

```{r Perform Tukey separation}
# Perform Tukey-adjusted pairwise comparisons with compact letter display
cld(treat_means, adjust = "tukey")
```
### Interpretation of question 4
Tukey’s separation showed that Treatment 2 resulted in significantly lower emergence than Treatment 1.

## Question 5
The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 

```{r Provided function}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

```

```{r Print plot from function}
print(plot_cldbars_onefactor(emergence_lm_simple, "Treatment"))
```

### Interpretation of question 5
Treatments that share the same letter are not significantly different from each other, while treatments that do not share letters are significantly different.Treatment 2 stands alone with a unique letter, meaning it has the lowest emergence, while several higher‑emergence treatments formed overlapping statistical groups, indicating a gradient rather than discrete treatment differences.