---
title: "Coding Challenge 4"
author: "Mary Fantasia"
date: "2026-02-19"
output: 
  md_document:
    variant: gfm
  html_document: 
   toc: true  
   toc_float: true  
  pdf_document: 
  word_document:  
---

# Question 1

### A YAMIL header displays the title of the Rmarkdown, the author of the document, the date it was last edited, and the output documents it will create.

the YAMIL header for the document looks like:
```{r eval=FALSE}
---
title: "Coding Challenge 4"
author: "Mary Fantasia"
date: "2026-02-19"
output: 
  md_document:
    variant: gfm
  html_document:
    toc: true
---
```

### Literate programming is a key player in reproducible code. 

includes annotating code, using programming settings to make the document easy to read

# Question 2

### Adding code to the document is easy, just add a code chunk

[Link to data](Noel, Z.A., Roze, L.V., Breunig, M., Trail, F. 2022. Endophytic fungi as promising biocontrol agent to protect wheat from Fusarium graminearum head blight. Plant Disease. https://doi.org/10.1094/PDIS-06-21-1253-RE)

### Code chunk 1

```{r Load packages and read data}
library (tidyverse)
library (ggplot2)
library(ggpubr)
library(ggrepel)

don.concentration <- read.csv("MycotoxinData.csv", na.strings="na")
```

### Code chunk 2

```{r Plot DON data}
don.treatment <- ggplot(don.concentration, aes(x = Treatment, y = DON, color = Cultivar, fill = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab ("DON (ppm)") +
  geom_point(position = position_jitterdodge()) +  
  geom_point(shape = 21, alpha = 0.6) +
  scale_fill_manual(values = c("#CC79A7", "#009E73")) +
  scale_color_manual(values = c("#000000", "#000000")) +
  facet_wrap(~Cultivar, scales = "free")  
don.treatment

don.concentration$Treatment <- as.factor(don.concentration$Treatment)
don.concentration$Treatment <- factor(don.concentration$Treatment, levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70"))
```

### Code chunk 3

```{r Plot 15ADON data}
x15adon.effect <- ggplot(don.concentration, aes(x = Treatment, y = X15ADON, color = Cultivar, fill = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab ("15ADON") +
  geom_point(position = position_jitterdodge()) +  
  geom_point(shape = 21, alpha = 0.6) +
  scale_fill_manual(values = c("#CC79A7", "#009E73")) +
  scale_color_manual(values = c("#000000", "#000000")) +
  facet_wrap(~Cultivar, scales = "free") 
x15adon.effect
```

### Code chunk 4

```{r Plot seed mass data}
seed.mass.effect <- ggplot(don.concentration, aes(x = Treatment, y = MassperSeed_mg, color = Cultivar, fill = Cultivar)) +
  geom_boxplot() +
  xlab("") +
  ylab ("Seed Mass (mg)") +
  geom_point(position = position_jitterdodge()) +  
  geom_point(shape = 21, alpha = 0.6) +
  scale_fill_manual(values = c("#CC79A7", "#009E73")) +
  scale_color_manual(values = c("#000000", "#000000")) +
  facet_wrap(~Cultivar, scales = "free") 
seed.mass.effect
```

### Code chunk 5

```{r Add t-test to all data}
# figure 1a
don.t.test <- don.treatment +
  geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
don.t.test

# figure 1b
x15adon.t.test <- x15adon.effect +
  geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
x15adon.t.test

# figure 1c
seed.mass.t.test <- seed.mass.effect +
  geom_pwc(aes(group = Treatment), method = "t_test", label = "p.adj.format")
seed.mass.t.test
```

### Code chunk 6

```{r Combining all plots into one}
figure1 <- ggarrange(
  don.t.test,
  x15adon.t.test,
  seed.mass.t.test,
  labels = "auto",
  nrow = 1, 
  ncol = 3,  
  common.legend = T
)

figure1
```





